\pagestyle{empty}

\vspace{10mm} 

\noindent Thank you.
\newpage

\vspace*{\fill}
I declare that I carried out this master thesis independently, and only with the cited
sources, literature and other professional sources.

I understand that my work relates to the rights and obligations under the Act No.
121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
University in Prague has the right to conclude a license agreement on the use of this
work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.

\vspace{10mm} 
\noindent In~Prague, \today\hspace{\fill}Jakub Marek\\
\newpage

\tableofcontents*
\newpage

\begin{otherlanguage}{czech}
\noindent
Název práce: Použití GPS v Monte Carlo lokalizaci\\
Autor: Jakub Marek\\
Katedra / Ústav: Katedra softwarového inženýrství\\
Vedoucí diplomové práce: RNDr. David Obdržálek, Katedra softwarového inženýrství\\

\noindent Abstrakt: V~této práci je popsán návrh a implementace
frameworku pro tvorbu řídících systémů pro mobilní roboty.
Jde o~jednoduchý modulární systém
usnadňující tvorbu řídících systémů a pokusy s~jejich jednotlivými moduly.
Součástí frameworku jsou také některé nástroje pro lokalizaci ve venkovním prostředí.

\noindent Klíčová slova: robotika, gps, pseudorange, lokalizace
\end{otherlanguage}

\vspace{25mm}

\noindent
Title: Outdoor robot localization\\
Author: Jakub Marek\\
Department / Institute: Department of Software Engineering\\
Supervisor of the master thesis: RNDr. David Obdržálek, Department of Software Engineering\\

\noindent Abstract:The work describes the design and implementation
of a framework for building control systems for mobile robots.
It is a simple modular system simplifying creation of control systems and experiments with the modules.
The framework contains also several tools for localization in outdoor environment.

\noindent Keywords: robotics, gps, pseudorange, localization

%TODO: Abstract!

\newpage

\pagestyle{plain}
\setcounter{page}{1}
